---
title: "R Notebook"
output: html_notebook
---

```{r}
# Load required packages
library(rethinking)
library(ggplot2)
library(dagitty)
```
# Problem 1
### DAG
```{r}
dag1 <- dagitty('dag {
    E [pos="0,1"]
    WL [pos="1,1"]
    CI [pos="1,0"]
    M [pos="0,0"]
    E -> WL
    E->CI->WL
    E<-M->CI
}')
plot(dag1)
```
### Identifying the variables to control for to estimate the direct effect of E on WL
```{r}
adjustmentSets(dag1, exposure='E', outcome='WL', effect = 'total')
```
We need to adjust for CI to estimate the direct effect of E on WL. An incorrect model would be to adjust for M instead, this would give the total effect of E on WL.

### Simulation
```{r}
# Function to simulate data and fit models
f1 <- function(n=100, bM_E=1, bM_CI=1, bE_CI=1, bCI_WL=1, bE_WL=1) {
  # Simulate data
  M <- rnorm(n)  # Motivation
  u <- rnorm(n)  # Random noise
  E <- rnorm(n, bM_E*M + u) # Regular Exercise
  CI <- rnorm(n, bM_CI*M + bE_CI*E + u) # Caloric Intake
  WL <- rnorm(n, bE_WL*E + bCI_WL*CI + u) # Weight loss
  
  # Fit models
  b_direct <- coef(lm(WL ~ E + CI))['E']  # Model controlling for CI, direct effect
  b_incorrect <- coef(lm(WL ~ E + M))['E']  # Model controlling for M, total effect
  
  return(c(b_direct, b_incorrect))  # Return coefficients
}
# Perform Monte Carlo simulation
sim1 <- mcreplicate(1e4, f1(), mc.cores = 8)
```
### Plotting the correct and incorrect versions of the estimate
```{r}
# Plot posterior distributions
range1 <- range(sim1[1,])
range2 <- range(sim1[2,])
xlim <- range(c(range1, range2))
ylim <- c(0,3.)

dens(sim1[1,], lwd=3, xlab='Coefficient of Exercise (Direct Effect)', xlim=xlim, ylim=ylim)
dens(sim1[2,], lwd=3, col=2, add=TRUE, xlim=xlim, ylim=ylim)
```
# Problem 2: Education Level and Job Satisfaction
### DAG
```{r}
dag2 <- dagitty('dag {
    EL [pos="0,1"]
    JS [pos="1,1"]
    F [pos="1,0"]
    PE [pos="0,0"]
    EL -> JS
    EL -> F -> JS
    EL <- PE -> JS
}')
plot(dag2)
```
### Identifying the variables to control for to estimate the total effect of EL on JS
```{r}
adjustmentSets(dag2, exposure='EL', outcome='JS', effect = 'total')
```
We need to adjust for both PE to estimate the total effect of EL on JS.

### Simulation
```{r}
# Function to simulate data and fit models
f2 <- function(n=100, bPE_EL=1, bPE_JS=1, bEL_F=1, bF_JS=1, bEL_JS=1) {
  # Simulate data
  PE <- rnorm(n)  # Motivation
  u <- rnorm(n)  # Random noise
  EL <- rnorm(n, bPE_EL*PE + u) # Regular Exercise
  F <- rnorm(n, bEL_F*EL + u) # Caloric Intake
  JS <- rnorm(n, bF_JS*F + bEL_JS*EL + u) # Weight loss
  
  # Fit models
  b_total <- coef(lm(JS ~ EL + PE))['EL']  # Model controlling for F and PE
  b_incorrect <- coef(lm(JS ~ EL))['EL']  # Model controlling for F only
  
  return(c(b_total, b_incorrect))  # Return coefficients
}
# Perform Monte Carlo simulation
sim2 <- mcreplicate(1e4, f2(), mc.cores = 8)
```
### Plotting the correct and incorrect versions of the estimate
```{r}
# Plot posterior distributions
range1 <- range(sim2[1,])
range2 <- range(sim2[2,])
xlim <- range(c(range1, range2))
ylim <- c(0,3.2)

dens(sim2[1,], lwd=3, xlab='Coefficient of Education Level (Total Effect)', xlim=xlim, ylim=ylim)
dens(sim2[2,], lwd=3, col=2, add=TRUE, xlim=xlim, ylim=ylim)
```

# Problem 3: Medication Use and Patient Recovery Rate
### DAG
```{r}
dag3 <- dagitty('dag {
    M [pos="0,1"]
    RR [pos="1,1"]
    SI [pos="1,0"]
    A [pos="0,0"]
    M -> RR
    M <- SI -> RR
    SI <- A -> RR
}')
plot(dag3)
```
### Identifying the variables to control for to estimate the direct effect of M on RR
```{r}
adjustmentSets(dag3, exposure='M', outcome='RR', effect = 'direct')
```
We need to adjust for SI to estimate the direct effect of M on RR

### Simulation
```{r}
# Function to simulate data and fit models
f3 <- function(n=100, bA_SI=1, bA_RR=1, bSI_M=1, bSI_RR=1, bM_RR=1.5) {
  # Simulate data
  A <- rnorm(n)  # Age
  u <- rnorm(n)  # Random noise
  SI <- rnorm(n, bA_SI*A + u)
  M <- rnorm(n, bSI_M + u)
  RR <- rnorm(n, bA_RR*A + bSI_RR*SI + bM_RR*M + u)
  
  # Fit models
  b_direct <- coef(lm(RR ~ M + SI))['M']  # Model controlling for SI
  b_incorrect <- coef(lm(RR ~ M))['M']  # Model not controlling for SI
  
  return(c(b_direct, b_incorrect))  # Return coefficients
}
# Perform Monte Carlo simulation
sim3 <- mcreplicate(1e4, f3(), mc.cores = 8)
```
### Plotting the correct and incorrect versions of the estimate
```{r}
# Plot posterior distributions
range1 <- range(sim3[1,])
range2 <- range(sim3[2,])
xlim <- range(c(range1, range2))
ylim <- c(0,4.3)

dens(sim3[1,], lwd=3, xlab='Coefficient of Medication (Direct Effect)', xlim=xlim, ylim=ylim)
dens(sim3[2,], lwd=3, col=2, add=TRUE, xlim=xlim, ylim=ylim)
```

# Problem 4: Advertising and Sales
### DAG
```{r}
dag4 <- dagitty('dag {
    A [pos="0,1"]
    S [pos="1,1"]
    EC [pos="1,0"]
    B [pos="0,0"]
    Q [pos="0.5,0.5"]
    A -> S
    B -> A
    B <- EC -> S
    S <- Q -> B
}')
plot(dag4)
```
### Identifying the variables to control for to estimate the total effect of EL on JS
```{r}
adjustmentSets(dag4, exposure='A', outcome='S', effect = 'total')
```
We need to either adjust for both EC and Q or adjust for B to estimate the total effect of advertising on sales.

# Problem 5
### DAG
```{r}
dag5 <- dagitty('dag {
    SM [pos="0,1"]
    MH [pos="1,1"]
    OSN [pos="1,0"]
    PI [pos="0,0"]
    SM -> MH
    SM <- OSN -> MH
    PI -> SM
}')
plot(dag5)
```
OSN is a fork

# Problem 6
### DAG
```{r}
dag6 <- dagitty('dag {
    PE [pos="0,1"]
    HC [pos="1,1"]
    D [pos="1,0"]
    US [pos="0,0"]
    PE -> HC
    PE <- D -> HC
    HC -> US
}')
plot(dag6)
```